Creating optimal conditions for reproducible data analysis in \R with `fertile'\protect\thanks{The authors gratefully acknowledge contributions from Hadley Wickham, Jenny Bryan, Greg Wilson, Edgar Ruiz, and other members of the \pkg{tidyverse} team.}

### 2020-10-11 STAT-paper-template/APA/response-to-reviewers.md

**Reviewer’s Comments:**

General:
*- the use of shims is likely unfamiliar to R users. Therefore, I have some concern about the ability of a user to control how aggressively this package performs. For example, if a user specifies an absolute path and receives an error message, it will be difficult for the user to determine what function was used to specify this error message and determine if there is a way to be less aggressive (to turn that into a warning, instead of an error, for example). My fear is that this functionality may limit use. However, I do like that the sandbox() function addresses this to an extent, allowing users to test things out and "get it right" before retuning to where the original error message was received.*

This is an excellent point! Part of the reason that fertile is so strict on this behavior is that in my own personal experience, many R coders (particularly those with little experience) tend to ignore warnings as long as their code executes properly. Having fertile’s path-related behavior be error based ensures that no one will skip over the messages that come up. However, my co-author and I do recognize that some users may still want to run these functions despite the warnings. In order to take this into account, we have added detail to the error messages produced by the shims. The new messages provide users with a method to override the error produced by fertile.

Specific minor suggestions (in order of manuscript):

- *A few suggested references to include in Intro and/or 2.1:* 
    - *reproducibility framework: Patil et al. (https://www.nature.com/articles/s41562-019-0629-z)*
    - *good enough practices: Wilson et al. (https://journals.plos.org/ploscompbiol/article?id=10.1371/journal.pcbi.1005510)*
    - *best practices: Wilson et al (https://journals.plos.org/plosbiology/article?id=10.1371/journal.pbio.1001745)*

@@

- *3.2 `list_checks()` is a nice function, but a table with the checks and a brief description of what that check does in the paper would be beneficial.*

This table has been added! Since list_checks() provides no additional benefit to the user than the information in this table, it is no longer mentioned.

- *in 3.1 and 3.2, it may be nice to describe explicitly whether it's possible to just use one or the other (proactive vs retroactive)....similar to how in 3.3 it's described how logging can be turned off (or reset). I could envision one wanting to develop without error and then check at the end (retroactive only). Philosophically, I could see that you would not want to build that in and that people could then just load your package at the end...but this may be something explicitly stated in the manuscript.*

Thank you for pointing this out. A comment to this effect has been added to the end of the description of retroactive behavior.

@@

- *Section 4: Suggestion to rename to future directions or something other than results, given the lack of results here (no fault of the authors, here! simply beyond one's control.) But, I did read and was bummed to see no results here after the section header was Results. Excited to see these results in the future, regardless!*

Agreed on this point. No results are discussed so that warrants a name change. Since the section is about an experiment, however, I have chosen to replace it with “Experimental Testing”. We also decided to run a second trial this semester, so future directions is no longer an applicable name as it is happening currently!

*Very minor* (feel free to ignore)
- *I haven't dug into the code fully so I may have missed it, but I could imagine that a user may want the ability to add their own shim to make a check for something as new packages/functions come out or for something specific to their own uses. Is there functionality for one to create/add their own shim to their installation of fertile?*

Thank you for this excellent idea. There is not currently a functionality for this, but it is definitely something of interest for the future. As a result, we have added it to the future work section.

### 2020-10-11 STAT-paper-template/APA/fertile.Rnw
added log example

\citet{cooper2017guide} focus on data analysis in \R and identify a similar list of important reproducibility components, reinforcing the need for clearly labeled, well-documented, and well-separated files. In addition, they recommend publishing a list of dependencies and using version control. 
\citet{broman} reiterates the need for clear naming and file separation while sharing several additional suggestions: keep the project contained in one directory, use relative paths, and include a \cmd{README}.
\textcolor{blue}{\citet{wilson2017good} argue that, to follow good practice, an analysis must be located in a well-organized directory, be supplemented by a file containing information about the project (i.e. a README), and a provide an explicit list of dependencies.}
\textcolor{blue}{{\citet{wilson2014best}} emphasize communication, recommending that code be human-readable and consistently styled for ease of understanding once shared.}

@@

\pkg{fertile} attempts to address these gaps in existing software by providing a simple, easy-to-learn reproducibility package that, rather than focusing intensely on a specific area \textcolor{blue}{(such as file structure)}, provides some information about a wide variety of aspects influencing reproducibility. \textcolor{blue}{For example, \pkg{fertile} addresses file structure in part but also considers the quality of project documentation, the use of file paths, and the behavior of random number generation}. \pkg{fertile} is flexible, \textcolor{red}{\st{offering benefits}} \textcolor{blue}{providing reports on reproducibility} to users at any stage in the data analysis workflow, and provides \R-specific features, \textcolor{blue}{such as tools to manage package dependencies}, which address certain aspects of reproducibility that can be missed by external project development software.

@@

\textcolor{blue}{Users who want to override this behavior and execute their function anyway are provided with a piece of code to execute that will achieve this for them.}

\pkg{fertile} is even more aggressive with functions (like \func{setwd}) that are almost certain to break reproducibility \textcolor{blue}{no matter the type of path passed to them}. \textcolor{red}{\st{, causing them to throw errors that prevent their execution and providing recommendations for better alternatives}}. \textcolor{blue}{When these functions are called, users are not presented with an option to override the resulting error. Instead, the only way to achieve the desired action is for the user to execute an alternative function that produces the same behavior but in a reproducible way.}

@@

description <- c(
  "Checks to make sure a single .Rproj file is found in the root directory of the project.",
  "Checks to make sure there are no nested .Rproj files in subfolders of the project.",
  "Looks for indication of a clear order in which to run R scripts.",
  "There are six different 'has_tidy_...' functions. Each checks to make sure that no files of its given type (images, code, data, etc.) are located in the root directory of the project",
  "Ensures that all files located in the project directory are being referenced or created by other files in the project.",
  "Checks to make sure there is a README file in the root directory.",
  "Checks to ensure that all code files contain an adequate number of comments (>10% of lines).",
  "Checks to make sure that no absolute paths are referenced in code.",
  "Checks to make sure that no paths in code lead outside the project directory.",
  "Checks to make sure code does not contain uncontrolled randomness.",
  "Examines code to determine whether it conforms to 'tidy' style guide recommendations.")

@@

Subsets of the sixteen tests can be invoked using the \pkg{tidyselect} helper functions \citep{R-tidyselect} in combination with the more limited \func{proj\_check\_some} function. \textcolor{blue}{These helper functions, which include \func{starts\_with}, \func{ends\_with}, \func{contains}, and \func{matches}, allow users to check for matches between a string or regular expression and a list of variables--in this case, searching for checks from the list provided by \func{list\_checks} whose names match a certain condition (e.g. ending with "root")}.

@@

\textcolor{blue}{Both the proactive and retroactive features available in \pkg{fertile} are enabled automatically upon the loading of the package. To prevent users from accidentally enabling or disabling features, \pkg{fertile} does not provide any option to control which features are active. However, those looking to only use the proactive features can do so by refraining from using the retroactive-specific functions while those looking to only use the retroactive features can wait to load \pkg{fertile} until they are ready to produce a report on their project.}

@@

<<read-csv, message=FALSE, include=FALSE>>=
library(purrr)
library(forcats)
read_csv(fs::path(proj_dir, "mice.csv"))
@

@@

\begin{enumerate}[noitemsep]
\item Identify an \R function that is likely to be involved in operations that may break reproducibility. Popular functions associated with only one package (e.g., \func{read\_csv} from \pkg{readr}) are ideal candidates.
\item Create a function in \pkg{fertile} with the same name that takes the same arguments \textcolor{red}{\st{(and always the dots ...)}}.
\item Write this new function so that it: a) captures any arguments, b) logs the name of the function called, c) performs any checks on these arguments, and d) calls the original function with the original arguments. Except where warranted, the execution looks the same to the user as if they were calling the original function. \textcolor{blue}{Since \pkg{fertile} only needs access to the file path argument, "file" is the only declared argument for most shims. The other arguments are managed using the dots (...), which capture and save them, allowing them to eventually be passed back to the original shimmed function (e.g. \func{readr::read\_csv})}.

@@

\textcolor{blue}{\pkg{fertile} also keeps track of dependencies in additional detail. Every time the code contained within a project is rendered by `fertile`, the package generates a file capturing the `sessionInfo()` just after the code was run. This file contains information about the R version in which the code was run, the list of packages that were loaded, and their specific versions.}

@@

\begin{itemize} [noitemsep]
  \item \textcolor{red}{\st{Expanding dependency management features to include R session information and package version numbers in addition to package names.}}
  \item \textcolor{red}{\st{Expanding code and documentation style features to analyze whether code has been properly commented in addition to checking for a README and tidy code style}}.
  \item Adding \cmd{make}-like functionality that can analyze an \R project structure and files and use this information to generate a Makefile. This Makefile would have information about target files and their prerequisites and would assist with making sure that re-running an analysis is done as quickly as possible by ensuring that only the necessary code and files that have been updated are run when rebuilding and re-running code.
  \item \textcolor{blue}{Adding a badge system to improve user experience and simplify the process of understanding how reproducible any given project is. Under this system, projects would receive badges from \pkg{fertile} based on the sets of reproducibility checks they pass. For example, projects that pass all of the checks on file paths would receive a badge on their project for "Reproducible File Paths" while those who pass checks involving file clutter and project structure might be awarded a badge for "Exellent File Organization."}
  \item \textcolor{blue}{Implementing a method through which users can add their own shims to the package. This would allow users to modify \pkg{fertile} to fit the functions they use most often. }
\end{itemize}

>>>>

\citet{cooper2017guide} focus on data analysis in \R and identify a similar list of important reproducibility components, reinforcing the need for clearly labeled, well-documented, and well-separated files. In addition, they recommend publishing a list of dependencies and using version control. 
\citet{broman} reiterates the need for clear naming and file separation while sharing several additional suggestions: keep the project contained in one directory, use relative paths, and include a \cmd{README}.
\textcolor{blue}{\citet{wilson2017good} argue that to follow good practice, an analysis must be located in a well-organized directory, be supplemented by a file containing information about the project (i.e., a README), and a provide an explicit list of dependencies.}
\textcolor{blue}{{\citet{wilson2014best}} emphasize communication, recommending that code be human-readable and consistently styled for ease of understanding once shared.}

@@

\pkg{fertile} attempts to address these gaps in existing software by providing a simple, easy-to-learn reproducibility package that, rather than focusing intensely on a specific area \textcolor{blue}{(such as file structure)}, provides some information about a wide variety of aspects influencing reproducibility. \textcolor{blue}{For example, \pkg{fertile} addresses file structure in part but also considers the quality of project documentation, the use of file paths, and the behavior of random number generation}. \pkg{fertile} is flexible, \textcolor{red}{\st{offering benefits}} \textcolor{blue}{providing reports on reproducibility} to users at any stage in the data analysis workflow.
It also provides \R-specific features, \textcolor{blue}{such as tools to manage package dependencies}, which address certain aspects of reproducibility that can be missed by external project development software.

@@

\textcolor{blue}{Users who want to override this behavior and execute their function anyway are provided with a snipped of code to execute that will achieve this for them.}

\pkg{fertile} is even more aggressive with functions (like \func{setwd}) that are almost certain to break reproducibility \textcolor{blue}{no matter the type of path passed to them}. \textcolor{red}{\st{, causing them to throw errors that prevent their execution and providing recommendations for better alternatives}}. \textcolor{blue}{When these functions are called, users are not presented with an option to override the resulting error. Instead, the only way to achieve the desired action is for the user to execute a different function that produces the same behavior but in a reproducible way.}

@@

description <- c(
  "Checks to make sure a single .Rproj file is found in the root directory of the project.",
  "Checks to make sure there are no nested .Rproj files in subfolders of the project.",
  "Looks for indication of a clear order in which to run R scripts.",
  "There are six different 'has_tidy_*()' functions. Each checks to make sure that no files of its given type (images, code, data, etc.) are located in the root directory of the project",
  "Ensures that all files located in the project directory are being referenced or created by other files in the project.",
  "Checks to make sure there is a README file in the root directory.",
  "Checks to ensure that all code files contain an adequate number of comments (>10% of lines).",
  "Checks to make sure that no absolute paths are referenced in code.",
  "Checks to make sure that no paths in code lead outside the project directory.",
  "Checks to make sure code does not contain uncontrolled randomness.",
  "Examines code to determine whether it conforms to 'tidy' style guide recommendations.")

@@

Subsets of the sixteen tests can be invoked using the \pkg{tidyselect} helper functions \citep{R-tidyselect} in combination with the more limited \func{proj\_check\_some} function. \textcolor{blue}{These helper functions, which include \func{starts\_with}, \func{ends\_with}, \func{contains}, and \func{matches}, allow users to check for matches between a string or regular expression and a list of variables. In this case, we search for checks whose names match a certain condition (e.g., contain the string "paths")}.

@@

\textcolor{blue}{Both the proactive and retroactive features available in \pkg{fertile} are enabled automatically upon the loading of the package. To prevent users from accidentally enabling or disabling features, \pkg{fertile} does not provide any option to control which features are active. However, those who want to use the proactive features can simply  refrain from using the retroactive-specific functions.
Conversely, those who want to use only the retroactive features can wait to load \pkg{fertile} until they are ready to produce a report on their project.}

@@

<<include = FALSE>>=
log_clear()
@

<<read-csv, message=FALSE, results = 'hide'>>=
library(purrr)
library(forcats)
read_csv(fs::path(proj_dir, "mice.csv"))
@

@@

\begin{enumerate}[noitemsep]
\item Identify an \R function that is likely to be involved in operations that may break reproducibility. Popular functions associated with only one package (e.g., \func{read\_csv} from \pkg{readr}) are ideal candidates.
\item Create a function in \pkg{fertile} with the same name that takes the same arguments \textcolor{red}{\st{(and always the dots ...)}}.
\item Write this new function so that it: a) captures any arguments, b) logs the name of the function called, c) performs any checks on these arguments, and d) calls the original function with the original arguments. Except where warranted, the execution looks the same to the user as if they were calling the original function. \textcolor{blue}{For most shims, "file" is the only argument of interest. The other arguments are managed using the dots (...), which capture and save them, allowing them to eventually be passed back to the original shimmed function (e.g. \func{readr::read\_csv})}.

@@

\textcolor{blue}{\pkg{fertile} also keeps track of dependencies in additional detail. Every time the code contained within a project is rendered by \pkg{fertile}, the package generates a file capturing the \func{sessionInfo} just after the code was run. This file contains information about the \R version in which the code was run, the list of packages that were loaded, and their specific versions.}

@@

\begin{itemize} [noitemsep]
  \item \textcolor{red}{\st{Expanding dependency management features to include R session information and package version numbers in addition to package names.}}
  \item \textcolor{red}{\st{Expanding code and documentation style features to analyze whether code has been properly commented in addition to checking for a README and tidy code style}}.
  \item Adding \cmd{make}-like functionality that can analyze an \R project structure and files and use this information to generate a Makefile. This Makefile would have information about target files and their prerequisites and would assist with making sure that re-running an analysis is done as quickly as possible by ensuring that only the necessary code and files that have been updated are run when rebuilding and re-running code.
  \item \textcolor{blue}{Adding a badge system to improve user experience and simplify the process of understanding how reproducible any given project is. Under this system, projects would receive badges from \pkg{fertile} based on the sets of reproducibility checks they pass. For example, projects that pass all of the checks on file paths would receive a badge on their project for "Reproducible File Paths" while those who pass checks involving file clutter and project structure might be awarded a badge for "Excellent File Organization."}
  \item \textcolor{blue}{Implementing a method through which users can add their own shims to the package. This would allow users to modify \pkg{fertile} to fit the functions they use most often. }
\end{itemize}
